---
title: "Frequently Asked Questions"
output:
  html_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```


## What does it mean when I run `testApp()` and it says, "Server did not update any output values within 3 seconds"?

The full message is this:

```
Server did not update any output values within 3 seconds. If this is expected, use wait_=FALSE, values_=FALSE, or increase the value of timeout_.
```

This happens when setting an input value does *not* result in an output changing.

Shinytest normally assumes that when you call `app$setInput()`, the setting of an input will also result in a changed output, so the test driver process will wait for an output to change before moving on to the next step in a test script. If no outputs change within the timeout period (which defaults to 3 seconds), it prints that message.

If, in your application, you expect that setting a particular input does not immediately result in any output changes, then you should call `app$setInput()` with `wait_=FALSE, values_=FALSE)`.

On the other hand, if you expect that setting the input should result in an output change, but the time exceeds the timeout, then you can call `app$setInput()` with, for example, `timeout_=10`.


## How can my app detect if it's running in shinytest?

Sometimes it can be useful to alter the behavior of your application when it's being tested. To detect when it's being tested, you can use `isTRUE(getOption("shiny.testmode"))`, as in:

```{r}
if (isTRUE(getOption("shiny.testmode"))) {
   # Do something special here
}
```

## Can I control the value shinytest uses for comparing test results?

The **shiny** package provides a way to preprocess both input (`shiny::snapshotPreprocessInput()`) and output (`shiny::snapshotPreprocessOutput()`) values before those values are used by shinytest for comparing test results. These functions can be helpful for avoiding comparison of possible very large objects and/or superfluous information that generates false positive test results. They can used directly in your shiny app used for testing and/or internally in R packages. For example, the [DT](https://github.com/cran/DT/blob/44fb07c/R/shiny.R#L114-L125) packages uses `shiny::snapshotPreprocessOutput()` to remove information that end-users need not concern themselves about. 
